---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# mOMOP

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/mOMOP)](https://CRAN.R-project.org/package=mOMOP)
<!-- badges: end -->

The goal of mOMOP is an ontology that crosswalk the [Minimal Common Oncology Data Elements (mCode)](https://mcodeinitiative.org/) with the OMOP Vocabularies. mCode and the 
mCode Initiative provides a common data language and model for data exchange 
between entities while the OMOP Vocabularies integrate data elements into the 
OMOP Common Data Model to serve analytics use cases.  

## Installation

You can install mOMOP with:

``` r
library(devtools)
install_github("meerapatelmd/mOMOP")
```

## mCode to OMOP

```{r}
library(mOMOP)
```


The crosswalk between mCode value sets and OMOP concepts can be accessed with the following pre-packaged dataframes: `CANCER_STAGING`, `GENOMICS`, `ICD10CM`, `LOINC`, `SNOMED`, `SPECIMEN`, `UNITS_OF_MEASUREMENT`.

### Cancer Staging

```{r example}
head(CANCER_STAGING)
```

### Genomics 

```{r}
head(GENOMICS)
```

### ICD-10-CM

```{r}
head(ICD10CM)
```

### LOINC 

```{r}
head(LOINC)
```

### SNOMED 

```{r}
head(SNOMED)
```

### Specimen 

```{r}
head(SPECIMEN)
```

### Units of Measurement 

```{r}
head(UNITS_OF_MEASUREMENT)
```

## Code of Conduct

Please note that the mOMOP project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
