---
title: "mCode to OMOP Standard Library"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{mCode to OMOP Standard Library}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = TRUE
)
```

```{r libraries}
library(mOMOP)
library(chariot)
library(tidyverse)
```


```{r connect}
conn <- chariot::connectAthena()
```

This vignette takes a look at the Value Sets in the mCode Data Dictionary to 
create the standard library of Oncology concepts, available for loading in this 
package.  

The mCode valuesets are retrieved from the publicly available data dictionary.

```{r valuesets}
value_sets <- get_value_sets()
value_sets
```

To map the valuesets to OMOP concepts, it is split by `Code System`, the 
equivalent to the `vocabulary_id` field in OMOP's Concept table, for joining. 
As of `r Sys.Date()`, there are 9 `Code Systems` used in mCode.  

```{r valuesets_by_vocab}
value_sets_by_vocab <- 
  value_sets %>%
  rubix::split_by(col = `Code System`)
names(value_sets_by_vocab)
```

## Specimen   

In mCode, Specimen representation is based on HL7's Code System while in OMOP,
the Specimen domain has its own subset of concepts. Furthermore, mCode's  
Specimen representation maps to other OMOP Concept Ids. The Specimen 
representation in this package therefore has 2 parts:  

1. A map between mCode and OMOP Concepts outside OMOP's Specimen domain  
1. OMOP's Specimen domain concepts  

### Map from mCode to OMOP  

Specimen representation is derived from the valuesets.  

```{r specimen}
specimen_library <- 
  value_sets_by_vocab$`http://terminology.hl7.org/CodeSystem/v2-0487` %>%
  rubix::format_colnames()
specimen_library
```

The specimen description is joined on the Concept Synonym Names in the 
Concept Synonym table to retrieve the Concept Id the mCode Specimen maps to.  

```{r specimen_omop_library1}
specimen_omop_library1 <-
join_on_concept_synonym_name(
  data = specimen_library,
  column = "code_description",
  case_insensitive = TRUE,
  conn = conn
) %>%
  select(-concept_synonym_name,
         -language_concept_id) %>%
  rename(match_concept_id = concept_id)
specimen_omop_library1
```

The complete Concept representation is taken from the Concept Id.  

```{r}
specimen_omop_library2 <-
  join_on_concept_id(
    data = specimen_omop_library1,
    column = "match_concept_id",
    conn = conn
  )
specimen_omop_library2
```

To maintain a one-to-one representation of the original grain of information, 
the OMOP mappings are pivoted on the OMOP Domain to see how each mCode Specimen 
maps to by domain.  

```{r}
specimen_omop_library3 <-
  specimen_omop_library2 %>%
  merge_strip(into = "concept",
              domain_id) %>%
  pivot_wider(id_cols =  !concept,
              names_from = domain_id,
              values_from = concept) %>%
  select(-match_concept_id) %>%
  distinct()
specimen_omop_library3
```

To gather a complete representation of Specimens, any missing Specimen concepts 
in OMOP are added to the library.

```{r}
omop_specimen <-
  chariot::queryAthena("SELECT *
                     FROM omop_vocabulary.concept 
                     WHERE domain_id = 'Specimen' AND concept_class_id = 'Specimen';",
                     conn = conn) %>%
  merge_strip(into = "Specimen") %>%
  select(-Specimen_id)
omop_specimen
```

```{r}
specimen_omop_library <-
specimen_omop_library3 %>%
  full_join(omop_specimen,
            by = "Specimen") %>%
  distinct()
specimen_omop_library
```

This file is written to the `data-raw/` folder for distribution if it does not 
already exist.  

```{r,}
file <- file.path(getwd(), "data-raw", "specimen.csv")
if (!file.exists(file)) {
  write_csv(x = specimen_omop_library,
            file = file)
}
```


## Cancer Staging   

mCode uses the AJCC TNM Staging system, which correlates with NCIt concepts in 
the OMOP vocabulary.  

```{r}
cancer_staging_library <-
  value_sets_by_vocab$`http://cancerstaging.org` %>%
  rubix::format_colnames()
cancer_staging_library
```


As it is represented in the OMOP Vocabulary, NCIt is not separated by Tumor, 
Node, and Metastasis like it is in mCode. 

```{r}
ncit_omop_library <-
  chariot::queryAthena("SELECT *
                     FROM omop_vocabulary.concept 
                     WHERE vocabulary_id = 'NCIt' AND concept_class_id = 'AJCC Category';",
                     conn = conn)
ncit_omop_library
```

NCIt `AJCC Category` concepts are therefore grouped based on pattern matching 
with the Concept Code.

```{r}
ncit_omop_library2 <- ncit_omop_library
ncit_omop_library2$value_set_name <- ""

ncit_omop_library2 <-
  ncit_omop_library2 %>%
  mutate(value_set_name = 
           case_when(
             grepl("^[cp]{1}M", concept_code) ~ "TNMDistantMetastasesCategoryVS",
             grepl("^[cp]{1}N", concept_code) ~ "TNMRegionalNodesCategoryVS",
             grepl("^[cp]{1}T", concept_code) ~ "TNMPrimaryTumorCategoryVS"
             )
         )
ncit_omop_library2
```

A correlate to mCode's `TNMStageGroupVS` value set is not present in the OMOP 
Vocabularies, but since it can be derived from the TNM categories, it is skipped.

```{r}
cancer_staging_omop_library <-
  cancer_staging_library %>%
  left_join(ncit_omop_library2,
            by = "value_set_name")
cancer_staging_omop_library
```

This data is written to a `cancer_staging.csv` in the `data-raw/` folder for 
distribution if it does not already exist.  

```{r,}
file <- file.path(getwd(), "data-raw", "cancer_staging.csv")
if (!file.exists(file)) {
  write_csv(x = cancer_staging_omop_library,
            file = file)
}
```


## Genomics  

HGVS, HGNC, and ClinVar are collapsed into a single `Genomics` category.  

### HGVS  

mCode uses HGVS (http://varnomen.hgvs.org), but these codes are not in the OMOP 
Vocabulary and cannot be derived from the website. Therefore, it is skipped for 
now. 

```{r}
hgvs_library <-
  value_sets_by_vocab$`http://varnomen.hgvs.org` %>%
  rubix::format_colnames()
hgvs_library
```

### HGNC  

Gene Names in mCode are presumed to be derived from HGNC and the entire HGNC 
subset of the OMOP Vocabularies are included in the library.   

```{r}
hgnc_library <-
  value_sets_by_vocab$`http://www.genenames.org/geneId` %>%
  rubix::format_colnames()
hgnc_library
```

```{r}
hgnc_omop_library1 <-
  chariot::queryAthena("SELECT *
                     FROM omop_vocabulary.concept 
                     WHERE vocabulary_id = 'HGNC';",
                     conn = conn)
hgnc_omop_library1
```

The OMOP HGNC concepts are joined to the mCode data dictionary set. 

```{r}
hgnc_omop_library2 <-
  hgnc_omop_library1 %>%
  mutate(value_set_name = "HGNCVS")
hgnc_omop_library2
```

```{r}
hgnc_omop_library <-
  hgnc_library %>%
  left_join(hgnc_omop_library2,
            by = "value_set_name")
hgnc_omop_library
```


### ClinVar  

Like HGNC, the entire ClinVar subset of the OMOP Vocabularies are included in 
the library.   

```{r}
clinvar_library <-
  value_sets_by_vocab$ClinVar %>%
  rubix::format_colnames()
clinvar_library
```


```{r}
clinvar_omop_library1 <-
chariot::queryAthena("SELECT *
                     FROM omop_vocabulary.concept 
                     WHERE vocabulary_id = 'ClinVar';",
                     conn = conn)
clinvar_omop_library1

clinvar_omop_library2 <-
  clinvar_omop_library1 %>%
  mutate(value_set_name = "ClinVarVS")
clinvar_omop_library2

clinvar_omop_library <-
  clinvar_library %>%
  left_join(clinvar_omop_library2,
            by = "value_set_name")
clinvar_omop_library
```

### Final Genomics Library  

The final genomics library file is written if it does not already exist. 
```{r}
genomics_omop_library <-
  bind_rows(hgnc_omop_library,
            clinvar_omop_library)

file <- file.path(getwd(), "data-raw", "genomics.csv")
if (!file.exists(file)) {
  write_csv(x = genomics_omop_library,
            file = file)
}

```


## Units of Measure  

```{r}
uom_library <-
  value_sets_by_vocab$`http://unitsofmeasure.org` %>%
  rubix::format_colnames()
uom_library
```


```{r}
ucum_omop_library <-
chariot::queryAthena("SELECT *
                     FROM omop_vocabulary.concept 
                     WHERE vocabulary_id = 'UCUM';",
                     conn = conn)
ucum_omop_library
```



## SNOMED  

The SNOMED subset of the library contains:  

* Explicitly stated concepts  
* Concepts that are descendants of a stated code  
* Concepts that are ancestors of a stated code  

To derive all concepts, the code in the `logical_definition` field is extracted 
based on the listed scenario above.  

```{r}
snomed_library <-
   value_sets_by_vocab$`SNOMED CT` %>%
  rubix::format_colnames() %>%
  distinct() %>%
  mutate_all(as.character) %>%
  extract(col = logical_definition,
          into = "descendants_of",
          regex = "includes codes descending from ([0-9]{1,})[^0-9]{1}.*$",
          remove = FALSE) %>%
  extract(col = logical_definition,
          into = "ancestors_of",
          regex = "excludes codes descending from ([0-9]{1,})[^0-9]{1}.*$",
          remove = FALSE) %>%
  mutate(all_codes = coalesce(code, ancestors_of, descendants_of)) %>%
  mutate_all(trimws)
```

```{r}
snomed_omop_library1 <-
chariot::join_on_concept_code(kind = "LEFT",
                              data = snomed_library,
                              column = "all_codes",
                              where_in_concept_field = "vocabulary_id",
                              where_in_concept_field_value = "SNOMED")
snomed_omop_library1 <-
  snomed_library %>%
  left_join(snomed_omop_library1,
             by = c("value_set_name", 
                    "ancestors_of",
                    "descendants_of",
                    "all_codes",
                    "code_system", 
                    "logical_definition", 
                    "code", 
                    "code_description")) %>%
  distinct()
snomed_omop_library1
```


```{r}
snomed_omop_library2a <-
  snomed_omop_library1 %>%
  filter(!is.na(descendants_of)) 

snomed_omop_library2a2 <-
  join_for_descendants(kind = "LEFT",
                       data = snomed_omop_library2a,
                       ancestor_id_column = "concept_id") %>%
  select(all_of(colnames(snomed_library)),
         starts_with("descendant_")) %>%
  rename_all(~ str_remove_all(., pattern = "^descendant_"))
snomed_omop_library2a2
```

```{r}
snomed_omop_library2b <-
  snomed_omop_library1 %>%
  filter(!is.na(ancestors_of))

snomed_omop_library2b2 <-
  join_for_ancestors(kind = "LEFT",
                     data = snomed_omop_library2b,
                     descendant_id_column = "concept_id") %>%
  filter(min_levels_of_separation != 0) %>%
  select(all_of(colnames(snomed_library)),
         starts_with("ancestor_")) %>%
  rename_all(~ str_remove_all(., pattern = "^ancestor_"))
snomed_omop_library2b2
```

```{r}
snomed_omop_library2c <-
  snomed_omop_library1 %>% 
  filter(is.na(descendants_of), is.na(ancestors_of))
snomed_omop_library2c
```

```{r}
snomed_omop_library <-
  bind_rows(snomed_omop_library2a2,
            snomed_omop_library2b2,
            snomed_omop_library2c)
snomed_omop_library
```


## LOINC  
```{r}
loinc_library <-
   value_sets_by_vocab$LOINC %>%
  rubix::format_colnames() %>%
  mutate_all(as.character) %>%
  mutate_all(trimws) %>%
  distinct()
loinc_library
```

```{r}
loinc_omop_library <-
chariot::join_on_concept_code(kind = "LEFT",
                              data = loinc_library,
                              column = "code",
                              where_in_concept_field = "vocabulary_id",
                              where_in_concept_field_value = "LOINC")
loinc_omop_library <-
  loinc_library %>%
  left_join(loinc_omop_library,
             by = c("value_set_name", 
                    "code_system", 
                    "logical_definition", 
                    "code", 
                    "code_description")) %>%
  distinct()
loinc_omop_library
```


## ICD-10 CM 

```{r}
icd10cm_library <-
   value_sets_by_vocab$`ICD-10-CM` %>%
  rubix::format_colnames() %>%
  mutate_all(as.character) %>%
  mutate_all(trimws) %>%
  mutate(code = str_replace_all(string = code,
                                pattern = "(^[A-Z]{1}[0-9A-Z]{2})([0-9A-Z]{1}.*$)",
                                replacement = "\\1.\\2")) %>%
  distinct()

icd10cm_library
```

```{r}
icd10cm_omop_library <-
chariot::join_on_concept_code(kind = "LEFT",
                              data = icd10cm_library,
                              column = "code",
                              where_in_concept_field = "vocabulary_id",
                              where_in_concept_field_value = "ICD10CM")

icd10cm_omop_library <-
  icd10cm_library %>%
  left_join(icd10cm_omop_library,
             by = c("value_set_name", 
                    "code_system", 
                    "logical_definition", 
                    "code", 
                    "code_description")) %>%
  distinct()

icd10cm_omop_library
```


```{r}
chariot::dcAthena(conn = conn)
```


# Missing Codes


