---
title: "mCode Valuesets to RDF"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{mCode Valuesets to RDF}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE
)
```

```{r setup}
library(tidyverse)
library(mOMOP)
```

The mappings between the mCode Value Sets and OMOP Concepts is converted into an 
RDF to view and edit in Protege.  

There are 3 vocabulary-level datasets in this package that require additional 
parsing for appropriate categorization into owl: `SNOMED`, `ICD10CM`, and `LOINC`. 
This categarization is done by assessing the source mCode value sets each belongs 
to.   

In mCode, `LOINC` is used only for tumor marker testing with no overlap with 
any other dataset. 

```{r}
unique(LOINC$value_set_name)
```

All the valuesets represented by `ICD10CM` overlaps with `SNOMED`.  

```{r}
unique(ICD10CM$value_set_name)
unique(SNOMED$value_set_name)
```

To manage the overlaps, `ICD10CM` and `SNOMED` are combined and split by value set. 

```{r}
ICD10CM_SNOMED <-
  bind_rows(ICD10CM,
            SNOMED) %>%
  rubix::split_by(col = value_set_name)
names(ICD10CM_SNOMED)
```

The names are adjusted to following the uppercase snake format for this project. 

```{r}
nms <- names(ICD10CM_SNOMED)
nms <- str_remove_all(nms, "VS$")
nms <- str_replace_all(nms, "([a-z]{1})([A-Z]{1})","\\1_\\2")
nms <- toupper(nms)
names(ICD10CM_SNOMED) <- nms
names(ICD10CM_SNOMED)
```

The other datasets are combined into a list with `LOINC` renamed to 
`TUMOR_MARKER_MEASUREMENT`. Each dataset in the `mcode_classes` list is now 
divided by "Class" for the OWL class hierarchy.

```{r}
other_value_sets <-
list(CANCER_STAGING = CANCER_STAGING, 
          GENOMICS = GENOMICS, 
          TUMOR_MARKER_MEASUREMENT = LOINC, 
          SPECIMEN = SPECIMEN, 
          UNITS_OF_MEASUREMENT = UNITS_OF_MEASUREMENT) 

mcode_classes <-
  c(ICD10CM_SNOMED,
    other_value_sets)

names(mcode_classes)
```

All the concepts in each class are converted to the `chariot` package's `strip` 
format to represent them as single instances in OWL. For the `SPECIMEN` class 
specifically, the original mCode `code` and `code description` are carried over 
as instances to provide the appropriate gap coverage in the specimen representation.

```{r}
mcode_classes2 <-
mcode_classes %>%
  map(mutate_all, as.character) %>%
  bind_rows(.id = "class") %>% 
  chariot::merge_strip(into = "concept") %>%
  # Specimen domain has additional columns
  unite(
    col = "mcode_concept", 
    code, code_description, 
    sep = " ", 
    remove = FALSE,
    na.rm = TRUE
    ) %>%
  mutate(concept = coalesce(concept, Specimen, mcode_concept)) %>%
  distinct() %>%
  rubix::split_by(col = "class")
```

The data is written to an Excel to load when calling the `data` function. 

```{r,message=FALSE}
file <- file.path(getwd(), "data-raw/mcode_rdf.xlsx")
if (!file.exists(file)) {
  openxlsx::write.xlsx(x = mcode_classes2,
                       file = file)
}
```

# Data Glossary 

```{r}
mcode_classes3 <-
  mcode_classes2 %>%
  map(select, concept)
mcode_classes3
```


